\stktokens{}: Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2018-07-10 fully-abstract-comp/popl2019/paper.tex
More work on section 5 and notes for discussion.

\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability-related instructions.
The standard instructions are: unconditional and conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
The simplest of the capability instructions simply inspect the properties of capabilities: type (\textt{gettype}), linearity (\texttt{getl}), range (\texttt{getb} and \texttt{gete}), current address or seal (\texttt{geta}) or permission (\texttt{getp}).
The current address (or seal) of a capability (or set of seals) can be shifted by an offset (\texttt{cca}) or set to the base address (\texttt{seta2b}).

@@

The reverse is possible using \texttt{splice}.
Sealables can be sealed using \texttt{cseal} and pairs of sealed capabilities can be unsealed by crossing security boundaries (\texttt{xjmp}, see below).

@@

% known stack base
% + more
At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
However, this does not make sure that the caller's stack frame is on top of the call stack.
The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they later can return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
At this point (Figure~\ref{fig:stack-base-abuse-b})), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
In \stktokens{}, the base address of the stack is fixed as a compile-time constant.
The caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to this fixed base address, as it was for the return token they gave to the callee.
In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.

% Where do we get the stack pointer from, how can we know it is linear
% The other attack.
% KJAA: It is unclear to me what 'at this point' refers to.
% At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
% However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
% In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
% To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
% The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
% The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
% Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
% At this point, the adversary has everything they need to return from either of the two callbacks.
% Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
% From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
% They are both correct as the adversarial component has created two call stacks, one for each of the components.
% In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
% We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
% A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
% This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.

In \stktokens{} we ensure that at the start of execution, the stack memory is only referenced by a single, linear stack capability.
Because of this, we can assert validity of the return token simply by checking its base address and splicing it with the caller's stack frame.
There is no need to check linearity, because we know that only linear capabilities to this memory exist.

@@

By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from stack frames on the call stack.
Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
In other words by inspection of the operational semantics, \srccm{} guarantees LSE and WBCF.
It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.

@@

\subsection{Logical relation}
\label{subsec:logical-relation}
The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
Unfortunately due to space considerations, we cannot give a comprehensive description of it.
Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensible description of a logical relation for a capability machine.
Finally, for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, worlds are used in a logical relation to model the memory of a language.
\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds should model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to some regions that depend on the kind of sub-world.
where the regions differ for each of the sub-worlds.
Generally, the regions consist of a world indexed relation that specifies the memories allowed by the region and a region type.
The region type indicates how the region can be addressed which is important for linearity.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
The $\spatial$ region may seem odd as it not addressable, but all the relations in the logical relation are world indexed.
If the register-file relation and the memory relation both have worlds with the same ownership, then they can both contain linear capabilities that address the same memory.
If a spatial region is owned by the register-file relation, then it still has to be modelled in the memory which it can use a spatial region for.
To capture this, we define a relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint. We also define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
In a sense, linear capabilities are revocable, so the sub-worlds that allow spatial regions must also have a $\revoked$ region to model revocation.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories ($\Wor$ given below):
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
\end{multline*}
\begin{multline*}
  \Regions = 
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup \\
    \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
    \{\revoked\}
\end{multline*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the memory can contain every kind of region because it is addressable by both normal and linear capabilities.
The call stack and free stack should only be addressable by linear capabilities, so they only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

The worlds $\World$ are defined as $(\Worldh \times \Worlds \times \Worldfs)$ and they are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them in the lemma below.
\begin{lemma}
  \label{thm:recursive-domain-eq}
  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ and there exists an isomorphism $\xi$ such that
  \[
    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
  \]
  and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
\end{lemma}
The future world relation $\future$ is fairly standard as it is extensional and requires regions to stay the same with the exception that $\spatial$ regions can become $\spatialo$ or $\revoked$ in a future world.

@@

  If $\npair{\stpair{C}{C}}\in\lrcomp(W_1)$ and $\npair{\stpair{P}{P}} \in \lrcomp(W_2)$, then
  $\npair[n']{(\plug{C_S}{P_S},\plug{C_T}{P_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.

@@

\begin{figure}
  \centering
  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
    % \draw[help lines,yellow] (0,0) grid (10,7);
    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };

@@

    \node[align=left] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
    \node[align=left] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };

@@

\section{Discussion}
\label{sec:discussion}
\begin{itemize}
\item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
\item Sharing stack references accross component boundaries is supported
\item Other notions of well-bracketedness (specifically one would be to allow different stacks)
\end{itemize}

\subsection{Practical applicability}
We believe there are good arguments for practical applicability of \stktokens{}.
The strong security guarantees are proven in a way that is reusable as part of a bigger proof of compiler security.
Its costs are
\begin{itemize}
\item a constant and limited amount of checks on every boundary crossing.
\item possibly a small memory overhead because every stack frame must be of non-zero length
\end{itemize}
The main caveat is that we rely on the assumption that capability machines like CHERI can be extended with linear capabilities in an efficient way.

Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like \texttt{move}, \texttt{store} and \texttt{load}.
Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
Our instructions are a bit clumsy from this point-of-view because, for example, \texttt{move} or \texttt{store} will zero the source register resp. memory location if the value being written is linear.
A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
Adding splice and split will also consume some opcode space.

Another problem is caused by the move semantics for \texttt{load} in the presence of multiple hardware threads.
In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
This means that a \texttt{load} of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
This is in principle not a problem except that a compare-and-swap is significantly slower than a regular \texttt{load} (on the order of 10x slower or more).
When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than \texttt{load}'s worst-case...
The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.

>>>>

\trgcm{}'s instruction set is somewhat basic with the instructions one expects on most low-level machine as well as capability machine specific instructions.
The instructions that can be considered standard are: unconditional and a conditional jump (\texttt{jmp} and \texttt{jnz}), copy between registers (\texttt{move}), instructions that load from memory and store to memory (\texttt{load} and \texttt{store}), and arithmetic operations (\texttt{plus}, \texttt{minus}, and \texttt{lt}).
The simplest of the capability machine specific instructions simply inspect the type of system native tokens or project one of their fields.
These are \texttt{gettype} (for getting the type of a system token) and \texttt{getp}, \texttt{getl}, \texttt{getb}, \texttt{gete}, and \texttt{geta} (for projecting permission, linearity, base, end, and current address(/seal), respectively).
There are two instructions for manipulating the current address(/seal) of a capability(/set of seals): \texttt{cca} changes the current address(/seal) of a capability(/set of seals) by some offset, and \texttt{seta2b} sets the current address(/seal) of a capability(/set of seals) to the base address(/seal).

@@

\trgcm{} also has an instruction that does the inverse of splitting, so a linear capability, or any capability for that matter, that has been split can be restored (\texttt{splice}).
\trgcm{} has instructions for sealing sealables (\texttt{cseal}) and unsealing pairs of sealed capabilities by crossing security boundaries (\texttt{xjmp}, see below).

@@

% known stack base
% + more
At this point, the caller checks that the return token is adjacent to the stack capability for the caller's local stack frame and they have the means to do so.
However, this does not make sure that the caller's stack frame is on top of the call stack.
The issue is that stack frames may not be tightly packed leaving space between stack frames in memory.
An adversarial callee may even intentionally leave a bit of space in memory above the caller's stack frame, so that they later can return out of order by returning the bit of the return token for the bit of memory left above the caller's stack frame.
This is illustrated in Figure~\ref{fig:stk-base-abuse}: In Figure~\ref{fig:stack-base-abuse-a}, a trusted caller has called an adversarial callee.
The adversary calls the trusted code back, but first they split the return token in two and store on the heap the part for the memory adjacent to the trusted caller's call frame (Figure~\ref{fig:stack-base-abuse-b}).
The trusted caller calls the adversary back using the precautions we have described so far (Figure~\ref{fig:stack-base-abuse-c}).
At this point (Figure~\ref{fig:stack-base-abuse-b})), the adversary has access to a partial return token adjacent to the trusted caller's first stack frame which allows the adversary to return from this call breaking WBCF.
For the caller to be sure that there are no hidden stack frames above its own, they need to make sure that the return token is exactly the same as the one they passed to the callee.
In \stktokens{}, the caller verifies the validity of the return token by checking whether the base address\footnote{The stack grows downwards in the address space} of a returned token corresponds to the base address of the return token they gave to the callee.
In the scenario we just sketched, the caller would be alerted to the attempt to break WBCF when the base address check of the return token fails in Figure~\ref{fig:stack-base-abuse-d}.

% Where do we get the stack pointer from, how can we know it is linear
% The other attack.
% KJAA: It is unclear to me what 'at this point' refers to.
At this point, the caller is able to verify whether a return token is spendable which allows them to decide whether they have the topmost stack frame on a call stack.
However, in order to have WBCF there should only be one call stack, so the caller should also be able to verify that they have the top stack frame on the call stack used by all well-behaved components.
In case a trusted component did not start the execution on the machine, i.e.\ they were called by another component, then the trusted component has no way to be sure whether the stack capability passed to them corresponds to the one and only call stack.
To illustrate how an adversarial component can break WBCF by using multiple different stack capabilities consider this example: An adversary calls a trusted component with a callback and some linear capability that they claim is the stack capability.
The trusted component executes using the stack capability given to them and at some point they invoke the callback, creating return tokens and so on.
The adversary stores the return pair and return token and calls another trusted component but with a new callback and a new linear capability that they claim is the stack capability.
Again, this trusted component also executes using the stack capability given to them and at some point they too invoke the callback, taking all the precautions we have described so far.
At this point, the adversary has everything they need to return from either of the two callbacks.
Specifically, they could return from the first callback invocation before they have returned from the second, breaking WBCF.
From a high-level perspective, the problem is that both trusted components will be able to verify that they have the top stack frame on the call stack.
They are both correct as the adversarial component has created two call stacks, one for each of the components.
In order to solve this problem, the well-behaved code needs to somehow agree on the call stack.
We do this by statically deciding on a global base address for the stack which is used for the entirety of the execution.
A caller makes sure that they are using the correct stack by checking that the base address of the stack capability they receive corresponds to the global stack base address.
This also means that on every return, the caller will check whether the return token's base address is equal to the global stack address.

Having a globally decided stack address also solves another challenge.
Namely, the stack capability has to be linear in order for a caller to use it as a token, so a caller has to make sure that a capability is in fact linear when it is proposed as the stack capability by an untrusted source.
This is, however, implicitly checked when the caller checks that the stack capability governs the global stack base address.
The reason is that the stack capability needs to start execution as a linear capability which means that the capability that governs the global stack base address to begin with is linear.
Due to the non-duplicability of linear capability and the fact that they cannot become non-linear during execution, the caller knows that if they have a capability that governs the global stack base address, then it must be linear.

@@

By inspecting the operational semantics of \srccm{}, we can see that it does not allow reads from inactive stack frames on the call stack.
Further, the built-in call and returns make sure that a stack frame is pushed on call, and a return is to the address in the topmost stack frame on the call stack.
In other words by inspection of the operational semantics, \srccm{} guarantees LSE and WBCF.
It is important to point this out as \stktokens{} guarantee exactly the LSE and WBCF we have on \srccm{}.

@@

\subsection{Logical relation}
\label{subsec:logical-relation}
The backbone of the full-abstraction proof is a cross language, step-indexed, Kripke logical relation.
\todo[inline]{write a bit more introduction that informally tells the purpose of this logical relation in relation to the full proof.}
Unfortunately due to space considerations, we do not give a comprehensive description of it.
Instead we highlight the parts specific to the logical relation we have used and refer to the technical report~\citep{technical_report} for the details and to~\citet{skorstengaard_reasoning_2017} for a more comprehensive description of a logical relation for a capability machine.
Finally for the sake of presentation, we will sometimes omit describing details, such as step indexing, that are important for the correctness but otherwise not interesting.

\subsubsection{Worlds}
\label{subsec:worlds}
Generally speaking, a world is used to model the memory of a language in a logical relation.
\srccm{} has three kinds of memory, the heap, the call stack, and the free stack, and our worlds model this by being triples of sub-worlds.
All of the sub-worlds are partial maps from region names $\RegionName$ (modelled as natural numbers) to regions where the regions differ for each of the sub-worlds based on what they model.
Generally, the regions consist of a region type and a world indexed relation that specifies the memories allowed by the region.
The region types models linearity as they indicate how the memory can be addressed.
We have three types of regions: $\pure$, $\spatial$, and $\spatialo$.
The $\pure$ regions can be addressed by normal capabilities, the $\spatialo$ regions by linear capabilities, and the $\spatial$ regions by nothing.
To prevent duplication of linear capabilities, two linear capabilities cannot address the same $\spatialo$, so ownership has to be distributed between worlds.
To this end, we have the relation $W_1 \oplus W_2$ that requires $W_1$ and $W_2$ to be the same except for the ownership which has to be disjoint.
%We define $W_1 \uplus W_2$ which joins together two worlds with disjoint regions.
The unaddressable $\spatial$ region makes sure that the memory it represents is modelled.
Linear capabilities can be seen as revocable which is modelled by the $\revoked$ region.
Finally, the $\pure$ region also contains a seal interpretation function.
The seal interpretation is a world-indexed function that specifies the sealables a seal can be used to seal.
We split the regions in two categories:
\begin{multline*}
  \Regionh = 
  \{\pure \} \times (\Wor \monnefun \URel{\MemSeg^2}) \times \\
  (\Seal \parfun \Wor \monnefun \URel{\SealableCaps \times \SealableCaps})
\end{multline*}
\begin{align*}
  \Regions ={} &
    \{\spatial \} \times (\Wor \monnefun \URel{\MemSeg^2}) \cup{} \\
    & \{\spatialo \} \times (\Wor \monnefun \URel{\MemSeg^2})\cup
    \{\revoked\}
\end{align*}
and define the sub-worlds as follows:
\begin{align*}
  \Worldh  &={} \RegionName \parfun (\Regions + \Regionh)\\
  \Worlds  &={} \RegionName \parfun (\Regions \times \Addr)\\
  \Worldfs &={} \RegionName \parfun \Regions
\end{align*}
The world for the heap can contain every kind of region because it can be addressed by both normal and linear capabilities.
The call stack and free stack is only addressable by linear capabilities, so their sub-worlds only allow regions from $\Regions$.
Regions in $\Worlds$ correspond to a stack frame, so they are paired with an address which models the return address.

Define worlds as $\World = (\Worldh \times \Worlds \times \Worldfs)$. They are recursive as they contain world indexed relations\footnote{We have cheated in the presentation and included the solution to the circularity already.}. Recursive worlds are common in Kripke models, and we use the method of \citet{Birkedal:2011:SKM:1926385.1926401,Birkedal_taste_2014} to construct them.
\begin{lemma}
  \label{thm:recursive-domain-eq}
  There exists a c.o.f.e.\ $\Wor$ and preorder $\future$ such that $(\Wor,\future)$ is a preordered c.o.f.e.\and there exists an isomorphism $\xi$ such that
  \[
    \xi : \Wor \cong \blater (\Worldh \times \Worlds \times \Worldfs)
  \]
  and for $\hat{W},\hat{W}' \in \Wor: \quad  \hat{W'} \future \hat{W}\text{ iff }\xi(\hat{W'}) \future \xi(\hat{W})$
\end{lemma}
The future world relation $\future$ is fairly standard.
It is extensional and requires regions to stay the same in future worlds except $\spatial$ regions can become $\spatialo$ or $\revoked$.

@@

  If $\npair{\stpair{\context}{\context}}\in\lrcomp(W_1)$ and $\npair{\stpair{\comp}{\comp}} \in \lrcomp(W_2)$, then
  $\npair[n']{(\plug{\context_S}{\comp_S},\plug{\context_T}{\comp_T})} \in \lrec(W_1 \uplus W_2)$, for all $n' < n$.

@@

\renewcommand{\comp}{C}
\begin{figure}
  \centering
  \begin{tikzpicture}[scale=0.8,every node/.style={scale=.9}]
    % \draw[help lines,yellow] (0,0) grid (10,7);
    \node at (5,4.7) { ${\src{\comp_1}\mathrel{\sconeq} \src{\comp_2}}$ };

@@

    \node[align=center] at (8.2,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
      $ {\src{\comp_2}} \cong {\src{\comp_2}}$};
    \node[align=center] at (1.8,3) { $ {\trg{\context}} \cong \trg{\context}$ \\
      $ {\src{\comp_1}} \cong {\src{\comp_1}}$};
    % \node at (9,2.7) { $e  {\src{C_1}} \cong \src{C_1}} : tau$ };
    % \node at (8.7,3.3) { $ {\trg{\context}} \cong \trg{\context} :{\emptyset},tau \ra e,{\cdots}$ };
    % \node at (.8,3) { $e  {\src{C_1}} \cong {\src{C_1}} : tau$ };

@@

\section{Discussion}
\label{sec:discussion}
% \begin{itemize}
% \item explain how fully abstract overlay semantics could form one pass of a verified secure compiler.
% \item Sharing stack references accross component boundaries is supported
% \item Other notions of well-bracketedness (specifically one would be to allow different stacks)
% \end{itemize}
\subsection{Full abstraction}
% - Full abstraction proofs difficult
% - Need to compile to a machine with enforcement mechanisms - capability machine an option
% - Full abstraction proofs modular, so other full abstraction proofs could target \srccm{} and thus have more abstractions to work with than if \trgcm{} was the target.

\subsection{Sharing the stack}
% ?
\subsection{Practical applicability}
We believe there are good arguments for practical applicability of \stktokens{}.
The strong security guarantees are proven in a way that is reusable as part of a bigger proof of compiler security.
Its costs are
\begin{itemize}
\item a constant and limited amount of checks on every boundary crossing.
\item possibly a small memory overhead because every stack frame must be of non-zero length
\end{itemize}
The main caveat is that we rely on the assumption that capability machines like CHERI can be extended with linear capabilities in an efficient way.

Although this assumption can only be discharged by demonstrating an actual implementation with efficiency measurements, the following notes are based on private discussions with people from the CHERI team, as well as our own thoughts on the matter.
As we understand it, the main problems to solve for adding linear capabilities to a capability machine like CHERI are related to the move semantics for instructions like $\tmove{}$, $\tstore{}$ and $\tload{}$.
Processor optimizations like pipelining and out-of-order execution rely on being able to accurately predict the registers and memory that an instruction will write to and read from.
Our instructions are a bit clumsy from this point-of-view because, for example, $\tmove{}$ or $\tstore{}$ will zero the source register resp. memory location if the value being written is linear.
A solution for this problem could be to add separate instructions for moving, storing and loading linear registers, at the cost of additional opcode space.
Adding splice and split will also consume some opcode space.

Another problem is caused by the move semantics for $\tload{}$ in the presence of multiple hardware threads.
In this setting, zeroing out the source memory location must happen atomically to avoid race conditions where two hardware threads end up reading the same linear capability to their registers.
This means that a $\tload{}$ of a linear capability will need to behave like an atomic operation, similar to a compare-and-swap instruction.
This is in principle not a problem except that a compare-and-swap is significantly slower than a regular $\tload{}$ (on the order of 10x slower or more).
When using \stktokens{}, loads of linear capabilities normally happen only when a thread has stored its return data capability on the stack and loads it back from there after a return.
Because the stack is a region of memory with very high thread affinity (no other hardware thread should access it, in principle), and which is accessed quite often, we are hopeful that well-engineered caching could reduce the high overhead of atomic loads of linear capabilities.
If such memory could be (usually) kept exclusively locked in a cache close to the processor, the overhead of atomic loads in \stktokens{} might be significantly less than $\tload$'s worst-case...
The processor could perhaps also (be told to) rely on the fact that race conditions should be impossible for loads from linear capabilities (which should in principle be non-aliased) and just use a non-atomic load in that case.
